\MakeUppercase{Graph Product Structure for Non-Minor-Closed Classes}

### 2022-11-04 k-planar.tex
Another typo fix
Fixed big hole in proof of Theorem 7

@@

  Suppose, without loss of generality, that $v_0$ participates in $x_i$.  Then $a(v_0)=x$ is a $T$-ancestor of $x_i$ which is a $T$-ancestor of $x_r$.  Finally $a(v_r)=y$ is a $T$-ancestor of $x_r$.  Therefore, both $x$ and $y$ are contained in the path from the root of $T$ to $x_r$, so at least one of $x$ or $y$ is a $T$-ancestor of the other.  This completes the proof of \cref{i-ancestor}
\end{proof}

@@

  Since $a\in C_x$, there exists a $T$-descendant $x'$ of $x$ such that $ax'\in E(J)$.  Therefore, there exists a shortcut $P:=v_0,\ldots,v_r$ in $\SS$ with $v_0\in S_{x'}$ and $v_r\in S_a$.  For each $i\in\{0,\ldots,r\}$, let $Y_{x_i}$ be the part of $\mathcal{Y}$ that contains $v_i$.  For each $i\in\{1,\ldots,r-1\}$, $v_i\in Y_{x_i}$ is an internal vertex of $P$, so $v_r$ participates in $x_i$.  Therefore $a=a(v_r)$ is a $T$-ancestor of $x_i$ for each $i\in\{1,\ldots,r-1\}$.

  Since $v_0,\ldots,v_r$ is a path in $G$, $x_0,\ldots,x_r$ is a lazy walk in $H$.\footnote{A \defin{lazy walk} is a graph $G$ is a walk in the pseudograph obtained by adding a self-loop at each vertex of $G$.}  Let $j$ be the minimum integer such that $x_j$ is a strict $T$-ancestor of $x$.  Since $x'=x_0$ is a $T$-descendant of $x$ and $a=x_r$ is a strict $T$-ancestor of $x$, $j\in\{1,\ldots,r\}$ and $x_{j-1}$ is a $T$-descendant of $x$ (possibly $x_{j-1}=x$).  Since $x_{j-1}x_j\in E(H)$, $x_j \in B_{x_{j-1}}$ and therefore $x_j\in B_x$.  Therefore $xx_j\in E(H^+)$.

  Therefore $W:=y_0,\ldots,y_s:=x,x_j,x_{j+1},\ldots,x_r$ is a lazy walk in $H^+$ of length $r\le k$.  For each $j\in \{0,\ldots,s\}$, let $z_j$ be the lowest common $T$-ancestor of $y_0,\ldots,y_j$.  The same reasoning used to conclude that $xx_j\in E(H^+)$ can be used to conclude that $z_{i-1}=z_i$ or $z_{i-1}z_i\in E(H^+)$ for each $i\in\{1,\ldots,s\}$.  Therefore $z_0,\ldots,z_s$ is a directed lazy walk in $\overrightarrow{H}^+$.  Finally, let $\overrightarrow{W}$ be the path in $\overrightarrow{H}^+$ obtained by removing removing duplicates from $z_0,\ldots,z_s$.  Then $\overrightarrow{W}$ is a directed path in $\overrightarrow{H}^+$ from $x$ to $a$ of length at most $s\le r\le k$.  This completes the proof of \cref{general-bag-size}.
\end{proof}

>>>>

\begin{clm}\label{short_path_in_h}
  For each $v\in V(G)$, $H$ contains a path $x_0,\ldots,x_r$ of length $r\le k-1$ with $v\in Y_{x_0}$ and $x_r:=a(v)$ such that $v$ participates in each of $x_0,\ldots,x_r$.
\end{clm}

\begin{proof}
  Let $x_0\in V(T)$ be the unique node with $v\in Y_{x_0}$ and let $x_r:=a(v)$.  If $x_0=x_r$ then the length-$0$ path with the single vertex $x_0=x_r$ satisfies the requirements of the lemma.  Assume now that $x_0\neq x_r$.

  Consider the subgraph $H_v:=H[x\in V(T):\text{$v$ participates in $x$}\}]$.  Since $v\in Y_{x_0}$, $v$ participates in $x_0$, so $x_0\in V(H_v)$.  We claim that, for each $x\in V(H_v)\setminus\{x_0\}$, $H_v$ contains a path $y_0,\ldots,y_s$ of length $s\le k-1$ from $y_0:=x_0$ to $y_s:=x$.  Indeed, since $v$ participates in $x\neq x_0$, $\mathcal{S}$ contains a shortcut $P$ with $v\in V(P)$ and that crosses $x$.  Therefore $P$ contains a path $v_0,\ldots,v_s$ from $v_0:=v$ to some $v_s\in Y_x$.   Furthermore $v_s$ is an internal vertex of $P$ and $P$ has length at most $k$, so $s\le k-1$.  For each $i\in\{0,\ldots,s\}$, let $Y_{y_i}$ be the part of $\mathcal{Y}$ that contains $v_i$.  For each $i\in\{1,\ldots,s\}$, the existence of $v_{i-1}v_{i}\in E(G)$ implies that $y_{i-1}=y_i$ or $y_{i-1}y_i\in E(H)$.
  Therefore $y_0,\ldots,y_s$ is a lazy walk\footnote{A \defin{lazy walk} in a graph $G$ is a walk in the pseudograph obtained by adding a self-loop at each vertex of $G$.} in $H_v$ of length at most $s\le k-1$.  This lazy walk contains a path from $y_0$ to $y_s$ of length at most $s\le k-1$, as promised.

  Since $H_v$ contains a path from $x_0$ to $x$ for each $x\in V(H_v)\setminus\{x_0\}$,  $H_v$ is connected.  Since $H_v$ is connected, $x_r$---which is the lowest common $T$-ancestor of $V(H_v)$---is itself a node of $V(H_v)$.  Therefore $H_v$ contains a path $x_0,\ldots,x_r$ of length most $k-1$.  By definition, $v$ participates in each of $x_0,\ldots,x_r$.
\end{proof}

@@

  Suppose, without loss of generality, that $v_0$ participates in $x_i$.  Then $a(v_0)=x$ is a $T$-ancestor of $x_i$ which is a $T$-ancestor of $x_r$.  Finally $a(v_r)=y$ is a $T$-ancestor of $x_r$.  Therefore, both $x$ and $y$ are contained in the path from the root of $T$ to $x_r$, so at least one of $x$ or $y$ is a $T$-ancestor of the other.  This completes the proof of \cref{i-ancestor}.
\end{proof}

@@

  Since $a\in C_x$, there exists a $T$-descendant $x'$ of $x$ such that $ax'\in E(J)$.  Therefore, there exists a shortcut $P:=v_0,\ldots,v_r$ in $\SS$ with $v_0\in S_{x'}$ and $v_r\in S_a$.  For each $i\in\{0,\ldots,r\}$, let $Y_{x_i}$ be the part of $\mathcal{Y}$ that contains $v_i$. Since $x_r\in S_{a}$, $a(v_r)=a$.  By \cref{short_path_in_h}, $H$ contains a path $y_0,\ldots,y_s$ from $y_0:=x_r$ to $y_s:=a$ of length at most $k-1$ such that $v_r$ participates in each of $y_0,\ldots,y_s$.  We will use this to show that $H$ contains a path $z_0,\ldots,z_{s'}$ from $z_0:=x$ to $z_{s'}:=a$ such that $v_r$ participates in each of $z_1,\ldots,z_{s'}$.   There are three cases to consider, depending on the location of $x_r$ relative to $x'$.

  \begin{enumerate}
    \item $x_r$ is a strict $T$-ancestor of $x$.  The shortcut $P=x_0,\ldots,x_r$ implies that $x_0$ participates in $x_{r-1}$.  Therefore $x'=a(x_0)$ is a $T$-ancestor of $x_{r-1}$, so $x$ is also a $T$-ancestor of $x_{r-1}$.  Since $x_{r-1}$ is a $T$-descendant of $x$ and $x_{r-1}x_r\in E(H)$, $x_r=y_0$ is contained in $B_x$. Therefore $xy_0$ is an edge of $H^+$, so the path $x,y_0,\ldots,y_s$ is a path in $H^+$ from $x$ to $a$ of length at most $s+1\le k$ and $v_r$ participates in each of $y_0,\ldots,y_s$.

    \item $x_r=x$.  In this case, $y_0,\ldots,y_s$ is a path from $x$ to $a$ of length at most $k-1$ and $v_r$ participates in each of $y_0,\ldots,y_s$.

    \item $x_r$ is a strict $T$-descendant of $x$.
      Since  $y_0$ is a strict $T$-descendant of $x$ and $y_s$ is a $T$-ancestor of $x$, at least one of $y_0,\ldots,y_s$, say $y_i$, is contained in $B_x$. Therefore $xy_i$ is an edge of $H^{+}$.  Therefore $x,y_i,\ldots,y_s$ is a path in $H^+$ from $x$ to $a$ of length at most $s+1\le k$ and $v_r$ participates in each of $y_i,\ldots,y_s$.

  \end{enumerate}
  In each case, we obtain a path $z_0,\ldots,z_{s'}$ of length at most $k$ from $z_0:=x$ to $z_{s'}:=a$ such that $v_r$ participates in each of $z_1,\ldots,z_{s'}$.  By definition, $a$ is a $T$-ancestor of $x$.  Since $v_r$ participates in each of $z_1,\ldots,z_{s'}$, $a=a(v_r)$ is also a $T$-ancestor of each of $z_1,\ldots,z_{s'}$.  Let $W$ be the sequence obtained from $z_0,\ldots,z_{s'}$ by replacing $z_i$ with the lowest common $T$-ancestor of $z_0,\ldots,z_i$, for each $i\in\{0,\ldots,s'\}$.  Then $W$ is a lazy walk in $H^+$ of length at most $k$ that begins at $z_0=x$ and ends at $z_{s'}=a$. By removing duplicates from $W$ we obtain a path in $H^+$ that is also a directed path in $\overrightarrow{H}^+$.  This completes the proof of \cref{general-bag-size}.
\end{proof}

### 2022-11-04 response2.tex
One small fix
Fixed big hole in proof of Theorem 7

@@

>>>>

\begin{response}
  Thanks to the referee for catching this embarrassing mistake.  In our previous round of simplifications, we oversimplified here.  The existence of the directed path in $\overrightarrow{H}^+$ does not come from the shortcut $v_0,\ldots,v_r$.  Rather, it comes from some shortcut that contains $v_r$ and some vertex in $Y_a$ (which must exist since $a=a(v_r)$ so $v_r\in Y_a$ or some shortcut that contains $v_r$ crosses $a$).  We have rewritten this part accordingly.
\end{response}

@@

\begin{response}
  We have replaced \\
  ``$P := (S_x : x \in V (T))$ is a sequence of pairwise disjoint sets that cover $V(G)$'' \\
  with \\
  ``$\mathcal{P}:=\{S_x : x\in V(T)\}$ is a partition of $V(G)$.''
\end{response}
